---
title: "Technical details"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Technical details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


In this article you can read about the calculations made to obtain the `wisselstroom` objects.

# Setup

```{r setup}
library(wisselstroom)
```

```{r}
# making the objects

# To read in your institution specific bekostigingsbestand, 
# change path_to_file accordingly.
path_to_file <- file.path(system.file("extdata", package = 'wisselstroom'),
                           "VLPBEK_2025_20240115_99XX.csv")

# read the data and place it in an R data.frame
my_vlpbek_data <- read_bek_data(path_to_file)

# make the my_vlpbek object from the my_vlpbek_data object
my_vlpbek <- vlpbek(my_vlpbek_data)


# make the my_vlpbek_compact object from the my_vlpbek object
my_vlpbek_compact <- compact_vlpbek(my_vlpbek)
```

# A closer look at the objects

## Data set as read in with `read_vlpbek()`

```{r, eval=FALSE}
my_vlpbek_data <- read_bek_data(path_to_file)
```

The first 10 rows and 9 columns:  

```{r}
my_vlpbek_data[1:10,1:9]
```

Have a look at the structure of the read-in example data:  

```{r}
str(my_vlpbek_data)
```

## vlpbek object as made with `vlpbek()`

```{r, eval = FALSE}
# make the actual my_vlpbek object from the my_vlpbek_data object
my_vlpbek <- vlpbek(my_vlpbek_data)
```

Have a look at the structure of the newly made `my_vlpbek` object:   

```{r}
str(my_vlpbek)
```

Compared to the original dataset, the VLP record is translated to first 3 items. The rows containing the SLR record (totals per record type) and the BLB records (studycareer totals per student) are ignored.  

All the BRD records (enrolment) are placed in its own `enrolment` `data.frame`, as are all the BRR records (degrees) in the `degrees` `data.frame`.  Only the relevant columns for wisselstroom insights are included. A column `academic_year` is added to both `data.frames`, calculated on basis of `date_enrolment` for the `enrolments`, and on `date_graduation` for the `degrees`. 

The column `student_id` is calculated from two original columns in the `my_vlpbek_data`, `V2` and `V3`:   
-  `V2` contains the BSN of the student
-  `V3` contains the onderwijsnummer (educationa number) of the student
-  if `V2` has a value, student_id gets that number , prefixed with a "b", short for BSN
-  if `V2` does not have a value, student_id becomes V3, with prefix "e" for educational number



## vlpbek_compact object as made with `compact_vlpbek()`

```{r, eval=FALSE}
# make the my_vlpbek_compact object from the my_vlpbek object
my_vlpbek_compact <- compact_vlpbek(my_vlpbek)
```


Have a look at the structure of `my_vlpbek` object:   

```{r}
str(my_vlpbek)
```

